{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-20 thesis/quicksand.tex
quicksand discussion ; also split dowefindbugsfaster into 2 figures

\Cref{fig:dowefindbugsfaster}
plots the bug-finding performance of Quicksand's three experimental trials
against the three control approaches
in a cumulative distribution of total bugs found against elapsed time.
The farthest-right point on each series indicates in how many total test cases that trial found a bug
after the 10 CPU-hour timeout.
\Cref{fig:dowefindbugsfaster}(a)
is the main, resource-fair comparison by CPU-time,
while \Cref{fig:dowefindbugsfaster}(b) shows
the wall-clock time until each bug was found.

\begin{figure}[p]
	\begin{tabular}{c}
        \includegraphics[width=0.71\textwidth]{../proposal/dowefindbugsfaster-v2.pdf} \\
		\begin{tabular}{p{\textwidth}}
                (a) Bugs found by Quicksand versus control experiments, measured in CPU time.
                %as a function of elapsed CPU time.
                Overall,
                a more resource-fair comparison than (b),
                although Quicksand's start-up overhead is exaggerated, as the SSS-MC tests are not parallelized. \\
		\end{tabular}
                \\
                \\
        \includegraphics[width=0.71\textwidth]{../proposal/dowefindbugsfaster-wallclock-v2.pdf} \\
                %(b) Bugs found by elapsed wall-clock time.
		\begin{tabular}{p{\textwidth}}
                (b) Bugs found by Quicksand versus control experiments, measured in wall-clock time.
                Quicksand is parallelized tenfold; the vertical line indicates its 1 hour limit. \\
		\end{tabular}
        \end{tabular}
	\caption[Bug-finding performance comparison
	between Quicksand and prior work.]
	{Bug-finding performance comparison between
        by several configurations of Quicksand and the single-state-space approaches.
	Quicksand finds 125\% as many bugs
        with data-race PPs
	at the 10-hour mark, compared to the best prior work approach.}
        \label{fig:dowefindbugsfaster}
\end{figure}

@@

Regarding Quicksand's tenfold parallelism,
before the break-even point at 200 seconds,
it lags behind SSS-MC-ICB due to the additional start-up overhead
of testing many state spaces at once even though the easy bugs may be found extremely quickly in any of them.
However, converting ICB's early CPU-time advantage
into faster wall-clock performance remains an open research problem \cite{parallel-dpor}.
\Cref{fig:dowefindbugsfaster}(b) gives Quicksand full credit for its inherent parallelism,
which ICB cannot yet practically match:
with a processor allocation of 10 CPUs, it outperforms all prior work approaches
for any fixed budget of wall-clock time
(i.e., comparing across a vertical line at $x=N$ for any $N$).
% After 1 hour of wall-clock time, tenfold Quicksand performs 158\% as well as SSS-MC-ICB. % totally meaningless

@@

\Cref{tab:drbugs} provides more detail on each of the bugs shown in \Cref{fig:dowefindbugsfaster},
broken down by test case.
The left half shows the number found by each experimental approach,
with the totals of each column corresponding to the values at $x=10$ hours in \Cref{fig:dowefindbugsfaster}.
In \mxtest, which checks the lock implementation for correctly providing mutual exclusion
(rather than trusting its correctness, as all other tests do),
SSS-MC-ICB and SSS-MC-DPOR
found dramatically fewer bugs (just 1).
Prior work has proposed {\em abstraction reduction} \cite{dbug-phdthesis},
in which verifying correctness properties of synchronization primitives
allows subsequently trusting them in other tests which use them to mitigate state space explosion;
\cref{sec:tm-abstraction} will explore this technique further.
By contrast, QS-Limited-HB found 10 mutex bugs, and SSS-MC-Shared-Mem found 12.
In the scope of this chapter,
this serves as strong evidence that new low-level synchronization code must be verified with data-race preemption points,
whether combined with Iterative Deepening or ICB.

\begin{table}[t]
	\begin{center}
	\small
	\begin{tabular}{r|c||c|c||c|c|c}
		%\multicolumn{2}{c||}{} & \multicolumn{5}{c||}{\bf {Total bugs}} & \multicolumn{4}{c||}{\bf {Data-race bugs}} \\
		% ---
		& {\bf Num.} & \multicolumn{2}{c||}{\bf Quicksand} & \multicolumn{3}{c}{\bf {Single-state-space MC}}
		%& \multicolumn{2}{c|}{\bf {Limited HB}} & \multicolumn{2}{c||}{\bf {Pure HB}} &
		%{\bf Mutual} & {\bf Avg. tested}
		\\
		% ---
		{\bf Test} & {\bf tested} & {\bf LHB} & {\bf PHB} & {\bf ICB} & {\bf DPOR} & {\bf ShMem}
		%& {\bf All} & {\bf Nondet.} & {\bf {All}} & {\bf {Nondet.}}
		%& {\bf timeouts} & {\bf subset SSes}
		\\
		% ---
		\hline
		%         #test qs-lhb qs-phb icb dpor every dronly nondets(LHB) dronly(PHB) nondets(LHB) mutual-TO comp.SSes
		{\tt broadcast\_test}    & 79  & 8   & 8   & 5   & 6   & 7   \\ % & 2 & 1 & {2} & {1} & 7    & 112.3 \\
		{\tt thr\_exit\_join}    & 79  & 23  & 20  & 13  & 13  & 14  \\ % & 11& 4 & {7} & {3} & {12} & {69.7} \\
		{\tt mutex\_test}        & 79  & 10  & 9   & 1   & 1   & 12  \\ % & 9 & 1 & {8} & {1} & 0    & -     \\
		{\tt paradise\_lost}     & 79  & 17  & 16  & 12  & 11  & 12  \\ % & 7 & 3 & {6} & {2} & {50} & {77.4} \\
		{\tt paraguay}           & 79  & 10  & 8   & 5   & 5   & 11  \\ % & 6 & 1 & {3} & {2} & 45   & {59.6} \\
		{\tt rwlock\_downgrade}  & 79  & 27  & 26  & 25  & 23  & 28  \\ % & 4 & 1 & {3} & {0} & {44} & {86.3} \\
		\hline
		{\tt priority-sema}      & 59  & 7   & 7   & 1   & 1   & 8   \\ % & 6 & 4 & {6} & {6} & 2    & 13.0  \\
		{\tt alarm-simultaneous} & 44  & 21  & 12  & 16  & 5   & 29  \\ % & 17& 1 & {7} & {6} & {17} & {7.8}  \\
		{\tt wait-simple}        & 52  & 30  & 26  & 24  & 23  & 1   \\ % & 7 & 2 & {2} & {0} & {15} & {33.8} \\
		\hline
		{\bf Total} & 629 & 153 & 132 & 102 & 88  & 122 \\ % 69& 15& {44}& {21}& {192}& {65.8} \\
	\end{tabular}
	\end{center}
	\caption[Summary of bugs found by each test program.]
	{Summary of bugs
	%and data races
	found by each test program.
	QS-LHB and QS-PHB are Quicksand; ICB/DPOR/ShMem are the controls (\cref{sec:quicksand-expt-setup}).
	}
	\label{tab:drbugs}
\end{table}

@@

\begin{table}[h]
	\begin{center}
		\small
	\begin{tabular}{r||c|c||c}
		{\bf Bound} & {\bf SSS-MC-ICB} & {\bf SSS-MC-Shared-Mem} & Prior work \cite{chess-icb} \\
		\hline
		0         & 2     & 1     & 3 \\
		1         & 82    & 86    & 7 \\
		2         & 16    & 32    & 5 \\
		3         & 2     & 3     & 1 \\
		4+        & 0     & 0     & 0 \\
		\hline
		\bf Total & 102   & 122   & 16 \\
	\end{tabular}
	\end{center}
	\caption[Distribution preemption bounds among bugs found by ICB.]
	{Distribution of preemption bounds among bugs found by ICB control experiments.
	Bound 0 means the bug was found by switching threads only on {\tt yield} calls.}
	\label{tab:icb-bounds}
\end{table}

@@

\Cref{fig:totalverif} plots the cumulative distribution of total verifications provided by each approach,
in the same style of graph as \Cref{fig:dowefindbugsfaster}.
For 167 of the 629 tests,
QS-Pure-HB was able to reach and complete the maximal state space with no bugs found,
hence providing the total verification guarantee justified by the proofs in \cref{sec:quicksand-soundness}.
QS-Limited-HB completed a verification for 153 of 629 tests,
slightly slower on account of Limited Happens-Before's higher false positive rate.
%(which Quicksand must confirm each time by running an extra instance of Landslide). % explained below
The next best approach for verifications was SSS-MC-Shared-Mem, which completed its search in only 39 cases.

\begin{figure}[t]
	\begin{center}
	\includegraphics[width=0.71\textwidth]{../proposal/totalverifs-v2.pdf}
	\end{center}
	\caption[Verification performance comparison
	between Quicksand and prior work.]
	{Verification performance comparison between
	Quicksand and single-state-space approaches.
	Among the latter, only SSS-MC-Shared-Mem is theoretically capable of verifying any test with data races;
	the others' series include only tests with no data races whatsoever,
	in which case synchronization preemption points alone suffice for a full verification.}
	\label{fig:totalverif}
\end{figure}

@@

When a model checking job times out,
the user would more likely prefer a summary of what parts of the test were verified
rather than to write off all the CPU time spent as wasted.
Prior work has proposed Probabilistic Concurrency Testing %(PCT)
\cite{randomized-scheduler},
which attempts to quantify the probability
that a bug may remain in some untested interleaving after a time-out,
although is opaque to the user about which parts of her code were actually tested.
Meanwhile, ICB proposes to inform the user
of the maximum number of preemptions that were used to test any individual interleaving,
under the assurance that most bugs are likely to be found with fewer preemptions (\cref{tab:icb-bounds}).
Quicksand instead reports which subsets of preemption points resulted in state spaces that did complete in time,
in hopes that the user can supplement such a result with her own intuition
by inspecting the code corresponding to the preemption points not tested (especially data races).
From prior work, Preemption Sealing \cite{sealing}
has argued the value of similar {\em compositional testing} when full verification is intractable,
deferring to the user's expertise to judge the value of each subset of preemption points verified.
\Cref{tab:partialverifs} shows Quicksand's partial verification results on timed-out tests.

\begin{table}[h]
	\begin{center}
	\small
	\begin{tabular}{r|c||c|c}
		% ---
		& {\bf Num.} & {\bf Mutual} & {\bf Avg. tested} \\
		{\bf Test} & {\bf tested} & {\bf timeouts} & {\bf subset SSes} \\
		% ---
		\hline
		%                        #test mutual-TO comp.SSes
		{\tt broadcast\_test}    & 79  & 7    & 112.3 \\
		{\tt thr\_exit\_join}    & 79  & {12} & {69.7} \\
		{\tt mutex\_test}        & 79  & 0    & -     \\
		{\tt paradise\_lost}     & 79  & {50} & {77.4} \\
		{\tt paraguay}           & 79  & 45   & {59.6} \\
		{\tt rwlock\_downgrade}  & 79  & {44} & {86.3} \\
		\hline
		{\tt priority-sema}      & 59  & 2    & 13.0  \\
		{\tt alarm-simultaneous} & 44  & {17} & {7.8}  \\
		{\tt wait-simple}        & 52  & {15} & {33.8} \\
		\hline
		{\bf Total}              & 629 & {192}& {65.8} \\
	\end{tabular}
	\end{center}
	\caption[Summary of partial verification results on timed-out tests.]
	{Summary of partial verification results on timed-out tests.
	``Mutual timeouts'' counts how often both QS-Limited-HB and SSS-MC-ICB
	(the best bug-finding approach from each group)
	timed out with no bug found.
	Among those, ``Average tested subset SSes'' counts how many partial verifications
	QS-Limited-HB provided on average for each test. %(\sect{\ref{sec:eval-sssmc}}). % ???
	}
	\label{tab:partialverifs}
\end{table}

@@

On the other hand, attempting to maximize the raw number of completed state spaces
may not be the most user-friendly way to present partial verification results.
For example, included in these numbers are small state spaces which are subsets of other state spaces also completed;
the user need not examine both subset and superset separately.
Future work could conduct user studies to determine the most effective way to present these partial results
from a software development perspective,
such as described in \cref{sec:future-friendly}.

@@

\begin{table}[t]
	\begin{center}
	\footnotesize
	\begin{tabular}{r|c||c|c|c|c||c|c|c||c}
		\multicolumn{2}{c||}{} & \multicolumn{4}{c||}{\bf {Data-race bugs}}
		& \multicolumn{3}{c||}{\bf {Verifications}} \\
		% ---
		& {\bf Num.}
		%& \multicolumn{2}{c|}{\bf Quicksand} & \multicolumn{3}{c||}{\bf {State-of-the-art}}
		& \multicolumn{2}{c|}{\bf {Limited HB}}
		& \multicolumn{2}{c||}{\bf {Pure HB}}
		& \multicolumn{3}{c||}{\bf {Pure HB}}
		& \multicolumn{1}{c}{\bf {Realloc.}}
		\\
		% ---
		{\bf Test} & {\bf tested}
		%& {\bf LHB} & {\bf PHB} & {\bf ICB} & {\bf DPOR} & {\bf ShMem}
		& {\bf All} & {\bf N.D.} & {\bf {All}} & {\bf {N.D.}}
		& {\bf DR PPs} & {\bf Benign} & {\bf Untested} & {\bf FPs} \\
		% ---
		\hline
		%                      #test dronly nond dronly nond
		%                                                   drpps benign untest FRMs
		{\tt broadcast\_test}    & 79  & 2 & 1 & {2} & {1} & 655  & 97  & 150 & 52  \\
		{\tt thr\_exit\_join}    & 79  & 11& 4 & {7} & {3} & 566  & 68  & 249 & 338 \\
		{\tt mutex\_test}        & 79  & 9 & 1 & {8} & {1} & 911  & 127 & 44  & 7   \\
		{\tt paradise\_lost}     & 79  & 7 & 3 & {6} & {2} & 783  & 2   & 414 & 166 \\
		{\tt paraguay}           & 79  & 6 & 1 & {3} & {2} & 936  & 9   & 510 & 180 \\
		{\tt rwlock\_downgrade}  & 79  & 4 & 1 & {3} & {0} & 543  & 1   & 310 & 156 \\
		\hline
		{\tt priority-sema}      & 59  & 6 & 4 & {6} & {6} & 65   & 51  & 3   & 0   \\
		{\tt alarm-simultaneous} & 44  & 17& 1 & {7} & {6} & 35   & 0   & 29  & 35  \\
		{\tt wait-simple}        & 52  & 7 & 2 & {2} & {0} & 71   & 1   & 28  & 31  \\
		\hline
		{\bf Total}              & 629 & 69& 15& {44}& {21}& 4565 & 356 & 1737& 965 \\
	\end{tabular}
	\end{center}
	\caption[Data race statistics among Quicksand experiments.]
	{Data race statistics among Quicksand experiments.
	``Data-race bugs'' counts, among Quicksand's bugs, how many required data-race preemption points to expose;
	among those, the ``N.D.'' (``nondeterministic'') columns show how many candidates
	required model checking to identify
	in the first place
	(\cref{sec:quicksand-eval-nondets}).
	%
	``Total DR PPs'' counts how many unique data-racing instructions
	QS-Pure-HB identified among tests where it found no bugs.
	Among those, ``Benign DRs'' counts how many were refuted as non-failing,
	while ``Untested DR PPs'' counts how many could not be checked in the time limit.
	Finally, ``Realloc. FPs'' counts how many reallocation false positives QS-Limited-HB suppressed.
	}
	\label{tab:drstatistix}
\end{table}

@@

\begin{figure}[t]
	\begin{center}
	\includegraphics[width=0.85\textwidth]{../proposal/nondets.pdf} % TODO: get some gf colors on here
	\end{center}
	\caption[Nondeterministic data-race preemption points are required to find some bugs.]
	{Nondeterministic data-race preemption points are required to find some bugs.
	Requiring stateless model checking integration to identify to begin with,
	incorporating these races as new preemption points
	allowed Quicksand to find
	% lhb: 127.77%
	% phb: 191.30%
	128\% (Limited HB) to 191\% (Pure HB) as many data-race bugs
	compared to using single-pass candidates alone.
	}
	\label{fig:dr-falsenegs}
\end{figure}

@@

% identifying & skipping redundancies in smaller ID state spaces via memoization (same applies to ICB!)

>>>>

\Cref{fig:dowefindbugsfaster-cpu}
plots the bug-finding performance of Quicksand's three experimental trials
against the three control approaches
in a cumulative distribution of total bugs found against elapsed CPU-time.
The farthest-right point on each series indicates in how many total test cases that trial found a bug
after the 10 CPU-hour timeout.
\Cref{fig:dowefindbugsfaster-wall} shows
the same experiments, measured by wall-clock time until each bug was found instead.

\begin{figure}[h]
	\begin{center}
        \includegraphics[width=0.9\textwidth]{../proposal/dowefindbugsfaster-v2.pdf} \\
	\end{center}
	\caption[Quicksand's bug-finding performance measured in CPU time.]
		{Quicksand's bug-finding performance measured in CPU time.
		Quicksand finds 125\% as many bugs with data-race preemption points at the 10-hour mark,
		compared to the best prior work approach.
		Quicksand's startup overhead is exaggerated, as the control experiments are not parallelized.}
        \label{fig:dowefindbugsfaster-cpu}
\end{figure}
% TODO: fix placement of the legend in this graph so that it's not awful page back and forthing
\begin{figure}[h]
	\begin{center}
        \includegraphics[width=0.9\textwidth]{../proposal/dowefindbugsfaster-wallclock-v2.pdf} \\
	\end{center}
	\caption[Quicksand's bug-finding performance measured in wall-clock time.]
		{Quicksand's bug-finding performance measured in wall-clock time.
		Quicksand is parallelized tenfold; the vertical line indicates its 1 hour limit.}
        \label{fig:dowefindbugsfaster-wall}
\end{figure}

%\begin{figure}[p]
%	\begin{tabular}{c}
%        \includegraphics[width=0.71\textwidth]{../proposal/dowefindbugsfaster-v2.pdf} \\
%		\begin{tabular}{p{\textwidth}}
%                (a) Bugs found by Quicksand versus control experiments, measured in CPU time.
%                %as a function of elapsed CPU time.
%                Overall,
%                a more resource-fair comparison than (b),
%                although Quicksand's start-up overhead is exaggerated, as the SSS-MC tests are not parallelized. \\
%		\end{tabular}
%                \\
%                \\
%        \includegraphics[width=0.71\textwidth]{../proposal/dowefindbugsfaster-wallclock-v2.pdf} \\
%                %(b) Bugs found by elapsed wall-clock time.
%		\begin{tabular}{p{\textwidth}}
%                (b) Bugs found by Quicksand versus control experiments, measured in wall-clock time.
%                Quicksand is parallelized tenfold; the vertical line indicates its 1 hour limit. \\
%		\end{tabular}
%        \end{tabular}
%	\caption[Bug-finding performance comparison
%	between Quicksand and prior work.]
%	{Bug-finding performance comparison between
%        by several configurations of Quicksand and the single-state-space approaches.
%	Quicksand finds 125\% as many bugs
%        with data-race PPs
%	at the 10-hour mark, compared to the best prior work approach.}
%        \label{fig:dowefindbugsfaster}
%\end{figure}

@@

Regarding Quicksand's tenfold parallelism,
before the break-even point at 200 seconds,
it lags behind SSS-MC-ICB due to the additional start-up overhead
of testing many state spaces at once even though the easy bugs may be found extremely quickly in any of them.
However, converting ICB's early CPU-time advantage
into faster wall-clock performance remains an open research problem \cite{parallel-dpor}.
\Cref{fig:dowefindbugsfaster-wall} gives Quicksand full credit for its inherent parallelism,
which ICB cannot yet practically match:
with a processor allocation of 10 CPUs, it outperforms all prior work approaches
for any fixed budget of wall-clock time
(i.e., comparing across a vertical line at $x=N$ for any $N$).
% After 1 hour of wall-clock time, tenfold Quicksand performs 158\% as well as SSS-MC-ICB. % totally meaningless

@@

\Cref{tab:drbugs} provides more detail on each of the bugs shown in \Cref{fig:dowefindbugsfaster-cpu},
broken down by test case.
The left half shows the number found by each experimental approach,
with the totals of each column corresponding to the values at $x=10$ hours in \Cref{fig:dowefindbugsfaster-cpu}.
In \mxtest, which checks the lock implementation for correctly providing mutual exclusion
(rather than trusting its correctness, as all other tests do),
SSS-MC-ICB and SSS-MC-DPOR
found dramatically fewer bugs (just 1).
Prior work has proposed {\em abstraction reduction} \cite{dbug-phdthesis},
in which verifying correctness properties of synchronization primitives
allows subsequently trusting them in other tests which use them to mitigate state space explosion;
\cref{sec:tm-abstraction} will explore this technique further.
By contrast, QS-Limited-HB found 10 mutex bugs, and SSS-MC-Shared-Mem found 12.
In the scope of this chapter,
this serves as strong evidence that new low-level synchronization code must be verified with data-race preemption points,
whether combined with Iterative Deepening or ICB.

\begin{table}[t]
	\begin{center}
	\small
	\begin{tabular}{r|c||c|c|c|c|c}
		%\multicolumn{2}{c||}{} & \multicolumn{5}{c||}{\bf {Total bugs}} & \multicolumn{4}{c||}{\bf {Data-race bugs}} \\
		% ---
		& {\bf Num.} & \multicolumn{2}{c|}{\bf Quicksand} & \multicolumn{3}{c}{\bf {Single-state-space MC}}
		%& \multicolumn{2}{c|}{\bf {Limited HB}} & \multicolumn{2}{c||}{\bf {Pure HB}} &
		%{\bf Mutual} & {\bf Avg. tested}
		\\
		% ---
		{\bf Test} & {\bf tested} & {\bf LHB} & {\bf PHB} & {\bf ICB} & {\bf DPOR} & {\bf ShMem}
		%& {\bf All} & {\bf Nondet.} & {\bf {All}} & {\bf {Nondet.}}
		%& {\bf timeouts} & {\bf subset SSes}
		\\
		% ---
		\hline
		%         #test qs-lhb qs-phb icb dpor every dronly nondets(LHB) dronly(PHB) nondets(LHB) mutual-TO comp.SSes
		{\tt broadcast\_test}    & 79  & 8   & 8   & 5   & 6   & 7   \\ % & 2 & 1 & {2} & {1} & 7    & 112.3 \\
		{\tt thr\_exit\_join}    & 79  & 23  & 20  & 13  & 13  & 14  \\ % & 11& 4 & {7} & {3} & {12} & {69.7} \\
		{\tt mutex\_test}        & 79  & 10  & 9   & 1   & 1   & 12  \\ % & 9 & 1 & {8} & {1} & 0    & -     \\
		{\tt paradise\_lost}     & 79  & 17  & 16  & 12  & 11  & 12  \\ % & 7 & 3 & {6} & {2} & {50} & {77.4} \\
		{\tt paraguay}           & 79  & 10  & 8   & 5   & 5   & 11  \\ % & 6 & 1 & {3} & {2} & 45   & {59.6} \\
		{\tt rwlock\_downgrade}  & 79  & 27  & 26  & 25  & 23  & 28  \\ % & 4 & 1 & {3} & {0} & {44} & {86.3} \\
		\hline
		{\tt priority-sema}      & 59  & 7   & 7   & 1   & 1   & 8   \\ % & 6 & 4 & {6} & {6} & 2    & 13.0  \\
		{\tt alarm-simultaneous} & 44  & 21  & 12  & 16  & 5   & 29  \\ % & 17& 1 & {7} & {6} & {17} & {7.8}  \\
		{\tt wait-simple}        & 52  & 30  & 26  & 24  & 23  & 1   \\ % & 7 & 2 & {2} & {0} & {15} & {33.8} \\
		\hline
		{\bf Total} & 629 & 153 & 132 & 102 & 88  & 122 \\ % 69& 15& {44}& {21}& {192}& {65.8} \\
	\end{tabular}
	\end{center}
	\caption[Summary of bugs found by each test program.]
	{Summary of bugs
	%and data races
	found by each test program.
	QS-LHB and QS-PHB are Quicksand; ICB/DPOR/ShMem are the controls (\cref{sec:quicksand-expt-setup}).
	}
	\label{tab:drbugs}
\end{table}

@@

\begin{table}[h]
	\begin{center}
		\small
	\begin{tabular}{r||c|c||c}
		{\bf Bound} & {\bf SSS-MC-ICB} & {\bf SSS-MC-Shared-Mem} & {\bf Prior work ICB} \cite{chess-icb} \\
		\hline
		0         & 2     & 1     & 3 \\
		1         & 82    & 86    & 7 \\
		2         & 16    & 32    & 5 \\
		3         & 2     & 3     & 1 \\
		4+        & 0     & 0     & 0 \\
		\hline
		\bf Total & 102   & 122   & 16 \\
	\end{tabular}
	\end{center}
	\caption[Distribution preemption bounds among bugs found by ICB.]
	{Distribution of preemption bounds among bugs found by ICB control experiments.
	Bound 0 means the bug was found by switching threads only on {\tt yield} calls.}
	\label{tab:icb-bounds}
\end{table}

@@

\Cref{fig:totalverif} plots the cumulative distribution of total verifications provided by each approach,
in the same style of graph as \Cref{fig:dowefindbugsfaster-cpu} and \Cref{fig:dowefindbugsfaster-wall}.
For 167 of the 629 tests,
QS-Pure-HB was able to reach and complete the maximal state space with no bugs found,
hence providing the total verification guarantee justified by the proofs in \cref{sec:quicksand-soundness}.
QS-Limited-HB completed a verification for 153 of 629 tests,
slightly slower on account of Limited Happens-Before's higher false positive rate.
%(which Quicksand must confirm each time by running an extra instance of Landslide). % explained below
The next best approach for verifications was SSS-MC-Shared-Mem, which completed its search in only 39 cases.

\begin{figure}[t]
	\begin{center}
	\includegraphics[width=0.9\textwidth]{../proposal/totalverifs-v2.pdf}
	\end{center}
	\caption[Verification performance comparison
	between Quicksand and prior work.]
	{Verification performance comparison between
	Quicksand and single-state-space approaches.
	Among the latter, only SSS-MC-Shared-Mem is theoretically capable of verifying any test with data races;
	the others' series include only tests with no data races whatsoever,
	in which case synchronization preemption points alone suffice for a full verification.}
	\label{fig:totalverif}
\end{figure}

@@

\subsubsection{Partial verification}
\label{sec:quicksand-eval-partial}

When a model checking job times out,
the user would more likely prefer a summary of what parts of the test were verified
rather than to write off all the CPU time spent as wasted.
To this end,
Quicksand
%instead
reports which subsets of preemption points resulted in state spaces that did complete in time,
in hopes that the user can supplement such a result with her own intuition
by inspecting the code corresponding to the preemption points not tested (especially data races).
From prior work, Preemption Sealing \cite{sealing}
has argued the value of similar {\em compositional testing} when full verification is intractable,
deferring to the user's expertise to judge the value of each subset of preemption points verified.
\Cref{tab:partialverifs} shows Quicksand's partial verification results on timed-out tests.

\begin{table}[h]
	\begin{center}
	\small
	\begin{tabular}{r|c||c|c}
		% ---
		& {\bf Num.} & {\bf Mutual} & {\bf Avg. tested} \\
		{\bf Test} & {\bf tested} & {\bf timeouts} & {\bf subset SSes} \\
		% ---
		\hline
		%                        #test mutual-TO comp.SSes
		{\tt broadcast\_test}    & 79  & 7    & 112.3 \\
		{\tt thr\_exit\_join}    & 79  & {12} & {69.7} \\
		{\tt mutex\_test}        & 79  & 0    & -     \\
		{\tt paradise\_lost}     & 79  & {50} & {77.4} \\
		{\tt paraguay}           & 79  & 45   & {59.6} \\
		{\tt rwlock\_downgrade}  & 79  & {44} & {86.3} \\
		\hline
		{\tt priority-sema}      & 59  & 2    & 13.0  \\
		{\tt alarm-simultaneous} & 44  & {17} & {7.8}  \\
		{\tt wait-simple}        & 52  & {15} & {33.8} \\
		\hline
		{\bf Total}              & 629 & {192}& {65.8} \\
	\end{tabular}
	\end{center}
	\caption[Summary of partial verification results on timed-out tests.]
	{Summary of partial verification results on timed-out tests.
	``Mutual timeouts'' counts how often both QS-Limited-HB and SSS-MC-ICB
	(the best bug-finding approach from each group)
	timed out. %with no bug found.
	Among those, ``Average tested subset SSes'' counts how many partial verifications
	QS-Limited-HB provided on average for each test. %(\sect{\ref{sec:eval-sssmc}}). % ???
	}
	\label{tab:partialverifs}
\end{table}

@@

@@

\begin{table}[t]
	\begin{center}
	\footnotesize
	\begin{tabular}{r|c||c|c|c|c||c|c|c||c}
		\multicolumn{2}{c||}{} & \multicolumn{4}{c||}{\bf {Data-race bugs}}
		& \multicolumn{3}{c||}{\bf {Verifications}} \\
		% ---
		& {\bf Num.}
		%& \multicolumn{2}{c|}{\bf Quicksand} & \multicolumn{3}{c||}{\bf {State-of-the-art}}
		& \multicolumn{2}{c|}{\bf {Limited HB}}
		& \multicolumn{2}{c||}{\bf {Pure HB}}
		& \multicolumn{3}{c||}{\bf {Pure HB}}
		& \multicolumn{1}{c}{\bf {Realloc.}}
		\\
		% ---
		{\bf Test} & {\bf tested}
		%& {\bf LHB} & {\bf PHB} & {\bf ICB} & {\bf DPOR} & {\bf ShMem}
		& {\bf All} & {\bf N.D.} & {\bf {All}} & {\bf {N.D.}}
		& {\bf DR PPs} & {\bf Benign} & {\bf Untested} & {\bf FPs} \\
		% ---
		\hline
		%                      #test dronly nond dronly nond
		%                                                   drpps benign untest FRMs
		{\tt broadcast\_test}    & 79  & 2 & 1 & {2} & {1} & 655  & 97  & 150 & 52  \\
		{\tt thr\_exit\_join}    & 79  & 11& 4 & {7} & {3} & 566  & 68  & 249 & 338 \\
		{\tt mutex\_test}        & 79  & 9 & 1 & {8} & {1} & 911  & 127 & 44  & 7   \\
		{\tt paradise\_lost}     & 79  & 7 & 3 & {6} & {2} & 783  & 2   & 414 & 166 \\
		{\tt paraguay}           & 79  & 6 & 1 & {3} & {2} & 936  & 9   & 510 & 180 \\
		{\tt rwlock\_downgrade}  & 79  & 4 & 1 & {3} & {0} & 543  & 1   & 310 & 156 \\
		\hline
		{\tt priority-sema}      & 59  & 6 & 4 & {6} & {6} & 65   & 51  & 3   & 0   \\
		{\tt alarm-simultaneous} & 44  & 17& 1 & {7} & {6} & 35   & 0   & 29  & 35  \\
		{\tt wait-simple}        & 52  & 7 & 2 & {2} & {0} & 71   & 1   & 28  & 31  \\
		\hline
		{\bf Total}              & 629 & 69& 15& {44}& {21}& 4565 & 356 & 1737& 965 \\
	\end{tabular}
	\end{center}
	\caption[Data race statistics among Quicksand experiments.]
	{Data race statistics among Quicksand experiments.
	``Data-race bugs'' counts, among Quicksand's bugs, how many required data-race preemption points to expose;
	among those, the ``N.D.'' (``nondeterministic'') columns show how many candidates
	required model checking to identify
	in the first place
	(\cref{sec:quicksand-eval-nondets}).
	%
	``Total DR PPs'' counts how many unique data-racing instructions
	QS-Pure-HB identified among tests where it found no bugs.
	Among those, ``Benign'' counts how many were refuted as non-failing,
	while ``Untested'' counts how many could not be checked in the time limit.
	Finally, ``Realloc. FPs'' counts how many reallocation false positives QS-Limited-HB suppressed.
	}
	\label{tab:drstatistix}
\end{table}

@@

\begin{figure}[t]
	\begin{center}
	\includegraphics[width=0.9\textwidth]{../proposal/nondets.pdf} % TODO: get some gf colors on here
	\end{center}
	\caption[Nondeterministic data-race preemption points are required to find some bugs.]
	{Nondeterministic data-race preemption points are required to find some bugs.
	Requiring stateless model checking integration to identify to begin with,
	incorporating these races as new preemption points
	allowed Quicksand to find
	% lhb: 127.77%
	% phb: 191.30%
	128\% (Limited HB) to 191\% (Pure HB) as many data-race bugs
	compared to using single-pass candidates alone.
	}
	\label{fig:dr-falsenegs}
\end{figure}

@@

This section will list the current limitations of Quicksand and Iterative Deepening
and discuss opportunities for future improvement.

\subsubsection{Avoiding redundant work}

When Quicksand extends a small state space with more preemption points,
the new state space is guaranteed to test a superset of interleavings compared to the old one.
Because Quicksand prioritizes completing small state spaces before their descendants,
the superset state spaces we run later will repeat each branch of their already-completed subsets,
and any interleaving which does not preempt threads on any of the new preemption points will be repeated work.
%
% lol i never did this
%We measured the proportion of repeated work among completed state spaces across our test suite;
%on average, {\bf \large 999\%} of the interleavings in each test were repeated, with some tests as high as {\bf \large 9999\%}.
This may make Quicksand slower than the single-state-space approach to find certain bugs,
for example, if both {\tt mutex\_lock()} and {\tt mutex\_unlock()} preemption points
together expose a bug, but not either alone.
Predicting whether an upcoming interleaving has already been tested is not straightforward,
but future implementations
of Iterative Deepening and/or ICB
could incorporate cross-job memoization
to prune some or all such repeated work.

Similarly, when pursuing total verification,
if the state space resulting from preempting on every instruction
(or equivalently, the maximal state space, thanks to \cref{sec:quicksand-soundness})
could be completed in time,
a model checker which immediately jumped to that state space,
abandoning all smaller subsets would certainly achieve verification faster.
Quicksand's maximal state space mode ({\tt -M}, see \cref{sec:quicksand-impl-modes})
can strike a middle ground between Iterative Deepening and single-state-space preempt-everywhere,%
\footnote{Thesis note: {\tt -M} mode was implemented after the conference paper's publication \cite{quicksand},
and will be evaluated alongside transactional memory later in \cref{sec:tm-eval}.}
but future implementations of Iterative Deepening could prioritize the maximal state space more flexibly still.
For example, pinning its job to one of the available processors regardless of the status of any smaller jobs
would avoid getting too flooded with smaller jobs to even begin the maximal job before time runs out.
When full verification is infeasible,
completing even an intermediate-sized job would allow immediately pruning all subset jobs thereof,
perhaps using a form of binary search (on the preemption point set size) to find an appropriately-sized intermediate job.

\subsubsection{Preemption point subsets}

Quicksand was able to partially guarantee safety for some preemption points
in 93\% of tests with too-large maximal state spaces (\cref{sec:quicksand-eval-partial}).
However, in 6 cases, no more than the minimal state space could be verified,
and in 18 others, no state spaces were completed at all.
Larger state spaces often result from finer-grained locking,
which can indicate a more intricate concurrent algorithm or an unnecessarily complicated design (or both).
Such programs may require even more rigorous verification than a program with a single global lock,
making them important to consider for future work.
While Quicksand uses {\tt within\_function} (\cref{sec:quicksand-impl-mc})
{\em statically} to restrict where preemption points could arise in advance of the test,
future
Iterative Deepening
implementations could use this mechanism to {\em dynamically} subset preemption points further,
making partial verification of larger tests possible,
potentially even involving the user with interface options
to enable and disable preemption points of her choosing at run-time.
\cref{sec:warpzone-heuristics} discusses this possibility further.

\subsubsection{Static data race analysis}

### 2018-12-20 thesis/tm.tex
quicksand discussion ; also split dowefindbugsfaster into 2 figures

In this section I review some of the evaluation's results in a broader context,
list the current limitations of Landslide's implementation,
and discuss open problems for future work.

>>>>

This section will review %some of
the evaluation's results in a broader context,
list the current limitations of Landslide's implementation,
and discuss open problems for future work.
